\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-09-06 smurf_paper.tex
numerous fiddles throughout

Submillimetre imaging cameras generally use bolometers to maximise
sensitivity, in which an absorber collects photons that pass through
broad-band filters, and the signal is read out using a thermometer
\citep[e.g.][]{griffin2002}. The sensitivity of bolometers in a camera
like \scuba\ are limited by white photon and phonon noise from the
instrument and ambient backgrounds. The low-frequency noise, however,
is dominated by sources which produce slow variations in mean backgrounds
(e.g., thermal variations within the cryostat, and the atmosphere for
ground-based cameras), and drifts in the readout electronics. Such
noise has a power spectrum $\propto 1/f^\alpha$ ($\alpha>0$), and the
frequency at which it is comparable to the white noise level is called
the ``$1/f$ knee''. Since the low-frequency noise is largely
correlated between all, or subsets, of the bolometers in time, it can
be suppressed during map-making, since astronomical signals have the
distinct property that they are fixed in a sky reference frame
(assuming they are not time-varying). If successful, the noise in the
resulting map is said to be ``white noise limited'', meaning that the
map noise is uncorrelated spatially, and has an amplitude that scales
as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD is the
noise-equivalent flux-density (the white noise level of a bolometer in
1\,s of integration), and $t$ is the amount of integration time in a
map pixel.

There are numerous ways to attack this map-making problem, both in
terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is sampled on a
range of temporal scales, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$ slew speeds), such that significant drift in
the bolometers due to low-frequency noise occurs more slowly than the
crossing times for the astronomical scales of interest; and (ii) by
offering scan patterns that cross the sky at a wide range of position
angles, ultimately observing each portion of the map at a number of
different times. Such methods are now used by virtually all existing
ground-based submillimetre cameras
\citep[e.g.][]{glenn1998,weferling2002,wilson2008,kovacs2008b}., in
favour of ``chopping'' methods (where the secondary is quickly moved
to modulate the signal) that were more appropriate for older
instruments that had poorer low-frequency noise performance, and were
only sensivitve to modest angular scales

There are three general styles of map-making that are relevant to
reducing bolomter data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the reduction of chopped data from
SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
\citep[e.g.][]{omont2001}, another camera from the same
generation. Generally speaking, such methods are fast, although
dependending on the science goals and noise properties of the data,
they may not achieve the best noise performance on the angular scales
of interest. In more recent years a method that has been popular for
reducing data in fields of faint point sources with BOLOCAM
\citep[e.g.][]{laurent2005}, and its younger sibling AzTEC
\citep[e.g.][]{scott2008}, is ``principal component analysis'' (PCA)
cleaning. A statistical black-box removes the most correlated
components of the bolometer signals, enabling the detection of
point-sources very close to the theoretical white-noise limits of the
detectors with reasonable computation times when small numbers of
bolometers are involved (hundreds rather than thousands of
detectors). However, PCA cleaning is not a good generic solution for
producing maps of extended structures, since such sources produce
correlated signals amongst many detectors, and are removed by this
procedure.

The best existing map-making strategies for recovering information on
all angular scales are \emph{maximum likelihood} techniques, in which
the time-series data are expressed as a sampling of the ``true'' map
of the sky plus noise, and then the equation is inverted to estimate
the map as some weighted linear combination of the data that minimizes
the variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the CMB literature include
\citet{tegmark1997} and \citet{stompor2002}, and an application to
data from SCUBA is described in \citet{borys2004}.  The downside to
such methods is that they can be both computationally expensive and
have large memory requirements. While for some experimental designs
fast iterative methods for the inversion do exist with minimal memory
usage \citep[e.g.][]{wright1996}, for the more general map-making
problem involving many detectors, the problem is significantly
complicated both by the need to measure the cross power-spectra for
all unique pairs of bolometers, and then to perform the inversion
itself. The most promising maximum-likelihood method that may one day
be applied to \scuba\ data is ``SANEPIC'', which successfully reduced
maps from the Balloon-borne Large-Aperture Submillimeter Telescope,
involving hundreds of bolometers, and correctly incorporating
inter-bolometer noise correlations \citep[][]{patanchon2008}.

%  Styles of map-making: least
%squares \citep{janssen1992} maximum likelihood
%\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
%implementation of \citet{wright1996} to remove chop,
%\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
%\citep[e.g.][]{laurent2005,scott2008,aguirre2010}.

@@

This paper is divided in the following way. We first describe the
properties of \scuba\ data and scan patterns in Section~\ref{sec:data}
to motivate the data model that we have adopted. Note that most of the
analysis described here applies equally to all of the data taken
to-date with \scuba, including during the Shared-Risks Observing
period (S2SRO) from 2010 February 11 to 2010 March 25 (during which
only single subarrays were available at each of the two observing
wavelengths). A comparison between the S2SRO and fully-commissioned
instrument is given in Section~\ref{sec:s2sro}. Next, the details of
the SMURF algorithm are given in Section~\ref{sec:algorithm}. The
paper is concluded in Section~\ref{sec:examples} with three detailed
test cases that span the majority of observation types likely to be
undertaken with \scuba, with an emphasis on the mitigation of
divergence problems, and characterizing the output maps: (i) Uranus, a
bright, compact source (Section~\ref{sec:point}); (ii) the Lockman
Hole, a blind survey of faint point-like source
(Section~\ref{sec:cosmo}); and (iii) the star-forming region M17,
including bright, extended emission (Section~\ref{sec:extended}). All
of the data analyzed in this paper are publicly available through the
CADC \scuba\ raw-data queries page for the dates and observation
numbers given in the
text\footnote{\texttt{http://www3.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/
jcmt/search/scuba2}}. All of the analysis was performed using
\texttt{Starlink} version ... .

@@

In this section we summarize how \scuba\ bolometers detect temperature
variations (Section~\ref{sec:bolos}), give exampels of typical
bolometers signals (Section~\ref{sec:bolosignal}), and the impact of
magnetic field pickup (Section~\ref{sec:magpickup}), and finally
illustrate the use of principal component analysis to explore the
correlated noise properties of \scuba\ data (Section~\ref{sec:pca}).

@@

Incoming light passes first through a beam-splitter, and then bandpass
filters, providing simultaneous illumination and wavelength definition
of both the 450 and 850\,\micron\ focal planes. The bolometers in each
focal plane consist of absorbers coupled to superconducting
transition-edge sensors (TESs) for thermometry. The focal planes are
held at a temperature somewhat below the critical temperatures of the
TESs by the fridge. The devices are then biased into the
superconducting transition using a combination of individual pixel
heaters (to compensate for most of the bulk variations in loading from
the atmosphere and other backgrounds), and an electrical voltage
bias. Throughout an observation the heater bias is fixed, but
smaller-scale load variations both from the backgrounds and, to a much
lesser extent, astronomical sources, significantly alter the TES
impedences as their temperatures vary. Thermoelectrical feedback then
keeps the TESs ``locked'' in the superconducting transition; an
increase in temperature increases the impedence, and lowers the
current, and therefore the heating resulting from the voltage
bias. These current changes produce varying magnetic fields that are
amplified using chains of superconducting quantum interference devices
(SQUIDs), and the larger output currents are digitized. Each detector
has its own SQUID for the first-stage of the amplification, but the
remaining stages occur within a common chain of SQUIDs for each of the
32 columns. All 40 rows are read-out in sequence at a row-visit
frequency of about 12\,kHz. Such a high sample rate is unnecessary to
produce maps, so the data are re-sampled to approximately 200\,Hz
before writing to disk.  It should also be noted that there is an
additional 41st row of SQUID readouts that are not connected to
TESs. These ``dark SQUIDs'' track non-thermal noise sources that are
common to each column's amplifier chain. The relationship between the
output digitized current, $I$, and the input power, $P$, or $dI/dP$ is
established using \emph{flatfield} observations immediately prior to
science observations, in which the output signal is measured
throughout a ramp of the pixel heaters (which provide a known input
power); see Section~2.1 in \citet{dempsey2012} and Section~5.3 in
\citet{holland2012}. Finally, the conversion to astronomical flux
units from power involves a correction for atmospheric extinction
\citep[primarily using the JCMT Water Vapour Monitor to track
line-of-sight opacity variations, see Section~3 in][]{dempsey2012},
and the application of a flux conversion factor (FCF) which is
established from regular measurements of calibrators such as Uranus
\citep[Section~5 in][]{dempsey2012}.

The predominant noise features in \scuba\ bolometers are: (i) the
effectively white thermal noise produced by the phonon and optical
loading on the absorbers (largely independent from bolometer to
bolometer); and (ii) longer timescale drifts resulting from slow
variations in the mean load (e.g. fridge, telescope, or atmospheric
changes), and also the readout electronics (often correlated amongst
many or all of the bolometers). As such, the power spectra of
bolometer time-series exhibit a ``$1/f^\alpha$'' ($\alpha>0$) shape
below some knee frequency, $f_\mathrm{knee}$, and they are then
relatively flat at higher frequencies. The spectra also include line
features which are predominantly produced by aliased high-frequency
noise sources detected during the multiplexed readout stage as
described above.  Finally, an anti-aliasing filter produces roll-off
in the spectrum at frequencies approaching Nyquist prior to the
200\,Hz re-sampling.

@@

In order to reduce the impact of noise on the final maps, it is
essential that an appropriate scan strategy be used. First, it is
clearly advantageous that the astronomical signals appear in the
lowest-noise region of the power spectrum, namely frequencies above
$f_\mathrm{knee}$. This is achieved by scanning the telescope at large
slew rates, which are practically limited to about
600\,arcsec\,s$^{-2}$ for the JCMT in order to have reasonable
turn-around overheads (the amount of time spent changing direction and
speed, over the largest elevation range possible). At this maximum
speed there is a sample every 3.0\,arcsec, or approximately one third
of the 450\,\micron\ diffraction-limited full-width at half-maximum
(FWHM) -- a typical rule-of-thumb for adequately sampling a Gaussian
point spread function (PSF). The scan \emph{patterns} are designed,
primarily, to mitigate the effects of low-frequency $1/f$ drift. The
simplest scan one might use is a raster ``boustrophedon'' pattern,
i.e. a series of parallel passes across the field. However, such a
scan yields extremely poor performace for structures that are
transverse to the scan direction, since a long time will have elapsed
between each parallel step. In addition, other periodic noise features
may conspire with such a regularly-spaced pattern to produce periodic
structures in the map, or \emph{scan-synchronous} noise.  It has
therefore been recognized that scan strategies that visit each region
of the map on as many different time scales, and scan angles as
possible, offer the best protection against large-scale noise
\citep[see discussion in][and references therein]{kovacs2008b}; such
strategies are said to offer good \emph{cross-linking}. As a tradeoff
between truly random scan patterns and feasibility of implementation
taking into consideration both mechanical and scheduling constraints,
\scuba\ presently has two primary scan modes: PONG, which covers
rectangular areas larger than the array footprint in a closed pattern
that ``bounces'' off the edges at 45\,degree angles (and also usually
combined with a rotation through a number of fixed position angles to
create a ``rotating PONG'' with even better cross-linking) which was
based on a similar scan mode developed at the Caltech Submillimeter
Observatory; and a ``constant velocity daisy'' (CV Daisy) for small
areas (of order the array footprint), in which the telescope moves in
a circle, whose centre also slowly traces out a circle. For a more
complete description of the \scuba\ observing modes, see Section~5 in
\citet{holland2012}.

\begin{figure*}
\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro.pdf}
\includegraphics[width=0.49\linewidth]{bolos_point_mix.pdf}
\caption{A comparison between single bolometer time series in each of
  the 450 and 850\,\micron\ bands with the mixing chamber temperature
  and azimuth/elevation pointing offsets, before and after upgrading
  the instrument. The grey signals overplotted in the top two panels
  show the residual time-series after removing the common-mode
  signals.  Left: data taken during the S2SRO period, observation 29
  on 20100313. There is a strong correlation between the bolometers
  and the roughly $\sim$25\,s oscillation in the fridge, but only a
  minor correlation with the telescope motion. The nearly flat
  common-mode subtracted signals show: (i) that most of the
  low-frequency signal is common to all of the bolometers; and (ii)
  the non-correlated, and predominantly white noise at 450\,\micron\
  is significantly larger than at 850\,\micron. Right: data taken with
  the fully-commissioned instrument, observation 38 on
  20111112. Unlike the S2SRO data, there is no strong signal produced
  by variations in the fridge temperature. The 450\,\micron\ data have
  minimal correlated low-frequency noise (as evidenced by the lack of
  a difference between the raw and common-mode subtracted signals),
  although the 850\,\micron\ data show a common signal correlated with
  the telescope motion (note that this scan has a significantly larger
  amplitude than the S2SRO data set), mostly likely caused by magnetic
  field pickup (see Section~\ref{sec:magpickup} and
  Fig.~\ref{fig:magpickup}).}
\label{fig:bolos_mix}
\end{figure*}

@@

In Fig.~\ref{fig:bolos_mix} we show sample time series from single
bolometers in each of the 450 and 850\,\micron\ focal planes, as well
as variations in the mixing chamber temperature (as a proxy for the
focal-plane temperature), and the telescope pointing, for two data
sets, before and after the upgrades that followed S2SRO.

@@

The low-frequency signal component of the S2SRO bolometer output is
also highly correlated amongst bolometers in the same subarray. We
have calculated a common-mode signal, $\mathbf{c}(t)$, as the average
time series of all the working bolometers. We then fit the amplitude
of $\mathbf{c}(t)$ at each wavelength to the signals shown in
Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey residual
signals. These residuals are quite flat, although still with
noticeable variations. The white noise is also apparent, and larger at
450\,\micron\ as one would expect from the larger backgrounds compared
to 850\,\micron.

@@

Next, we produce power spectral density (PSD) plots for four of the
most sensitive bolometers from both focal planes, using the same two
data sets, in Fig.~\ref{fig:pspec}. To produce this figure, we follow
the convention that the PSD as a function of frequency,
$\mathbf{P}(f)$, is normalized such that the integral over frequency
gives the same variance as the time-series variance across the full
time series. In other words, given a bolometer signal $\mathbf{b}(t)$,
%
\begin{equation}
\label{eq:psd}
\langle\mathbf{b}^2(t)\rangle = 2 \int_0^{f_\mathrm{N}} \mathbf{P}(f)
df ,
\end{equation}
%
where we only integrate over the positive frequencies up to
$f_\mathrm{N}$, the Nyquist frequency, and the factor of 2 accounts
for the (equal) power that appears at negative frequencies in the
discrete Fourier Transform. The units of the PSD written in this form
are pW$^2$\,Hz$^{-1}$.

@@

An additional noise source that is significant in only a small subset
of the data (and moreso during the S2SRO period) is magnetic field
pickup. Since the bolometer signals are ultimately detected through
the amplification of magnetic fields, any additional changing fields
within the cryostat will also be detected.

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{pca.pdf}
\caption{The first twelve components from a principal component
  analysis, ranked by the mean eigenvalues, of combined 450 (s4a
  subarray) and 850\,\micron\ (s8b subarray) time-series bolometer
  data, for the same observation as that used in the right-hand panels
  of Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec}. For each component
  the top plot shows the time series of the normalized eigenvectors,
  the middle plot its PSD, and the bottom coloured panels the
  eigenvalues for the bolometers at each wavelength (the amplitudes of
  the eigenvectors in the time series). For reference, both the mean,
  $\bar{\lambda}$, and \rms, $\lambda_\mathrm{rms}$, eigenvalues for
  the bolometers in each subarray are also shown. Most of the
  correlated $1/f$ signal is encapsulated in Components~1 and 2, and
  appears to be dominated by scan-synchronous noise (compare with the
  scan pattern in the right-hand panel of Fig.~\ref{fig:bolos_mix}).
  Components~3, 4, 7 and 11 are transient glitches that only appear in
  a small subset of the bolometers. Component~5 also appears in only a
  few bolometers, and is dominated by a narrow line at 57\,Hz which
  suggests that it is related to the 60\,Hz mains. Components~6 and
  8--10 contain a mixture of broad lines above $\sim$30\,Hz (which are
  probably aliased high-frequency noise), with some low-frequency
  drifts. Finally, the erratic appearance of the eigenvector for
  Component~12 is typical of a poorly-biased detector.}
\label{fig:pca}
\end{figure*}

@@

The majority of the correlated signal at both wavelengths in
Fig.~\ref{fig:pca} is clearly produced by Components~1 and 2. In both
cases, the eigenvector time series exhibit a roughly periodic signal
that resembles the scan pattern in the right-hand panel of
Fig.~\ref{fig:pca}. While these correlated signals also clearly appear
at both wavelengths, referring to the maps of eigenvalues, they are
considerably stronger at 850\,\micron\ (consistent with the visual
appearance of the bolometer signals in Fig.~\ref{fig:pca}). Also note
that the PSDs for these eigenvectors exhibit nearly a pure $1/f$
signature with almost no line features, or a white-noise plateau
(suggesting that these are high-\snr\ measurements of a purely
low-frequency drift).

@@

Detecting and correcting such a wide variety of steps reliably has proved
to be a challenge. In outline, the following stages
are involved in detecting steps in a single bolometer time stream:

@@

\item median smooth the whole time stream;

\item find the gradient of the median smoothed time series at each
sample;

>>>>

Submillimetre imaging cameras generally use bolometers, rather than
coherent detectors, to maximise sensitivity. The sensitivity of
bolometers is limited by white photon and phonon noise from the
instrument and ambient backgrounds. The low-frequency noise, however,
is typically dominated by sources which produce slow variations the
background (e.g., thermal variations within the cryostat, and the
atmosphere for ground-based cameras), and drifts in the readout
electronics. Such noise has a power spectrum $\propto 1/f^\alpha$
($\alpha>0$), and the frequency at which it is comparable to the white
noise level is called the ``$1/f$ knee''. Since the low-frequency
noise is largely correlated between all, or subsets, of the bolometers
in time, it can be suppressed during map-making, since astronomical
signals have the distinct property that they are fixed in a sky
reference frame (assuming they are not time-varying). If successful,
the noise in the resulting map is said to be ``white noise limited'',
meaning that the map noise is uncorrelated spatially, and has an
amplitude that scales as the $\mathrm{NEFD}/\sqrt{t}$, where the NEFD
is the noise-equivalent flux-density (the white noise level of a
bolometer in 1\,s of integration), and $t$ is the amount of
integration time in a map pixel.

There are numerous ways to attack this map-making problem, both in
terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is sampled on a
range of temporal scales, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$ slew speeds), such that significant drift in
the bolometers due to low-frequency noise occurs more slowly than the
crossing times for the astronomical scales of interest; and (ii) by
offering scan patterns that cross the sky at a wide range of position
angles, ultimately observing each portion of the map at a number of
different times. Such methods are now used by virtually all existing
ground-based submillimetre cameras
\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}., in
favour of ``chopping'' methods (where the secondary is quickly moved
to modulate the signal) that were more appropriate for older
instruments that had poorer low-frequency noise performance, and were
only sensivitve to modest angular scales

There are three general styles of map-making that are relevant to
reducing bolomter data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the reduction of chopped data from
SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
\citep[e.g.,][]{omont2001}, another camera from the same
generation. Generally speaking, such methods are fast, although
dependending on the science goals and noise properties of the data,
they may not achieve the best noise performance on the angular scales
of interest. In more recent years a method that has been popular for
reducing data in fields of faint point sources with BOLOCAM
\citep[e.g.,][]{laurent2005}, and its younger sibling AzTEC
\citep[e.g.,][]{scott2008}, is ``principal component analysis'' (PCA)
cleaning. A statistical black-box removes the most correlated
components of the bolometer signals, enabling the detection of
point-sources very close to the theoretical white-noise limits of the
detectors with reasonable computation times when small numbers of
bolometers are involved (hundreds rather than thousands of
detectors). However, PCA cleaning is not a good generic solution for
producing maps of extended structures, since such sources produce
correlated signals amongst many detectors, and are removed by this
procedure.

The best existing map-making strategies for recovering information on
all angular scales are \emph{maximum likelihood} techniques, in which
the time-series data are expressed as a sampling of the ``true'' map
of the sky plus noise, and then the equation is inverted to estimate
the map as some weighted linear combination of the data that minimizes
the variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the CMB literature include
\citet{tegmark1997} and \citet{stompor2002}, and an application to
data from SCUBA is described in \citet{borys2004}.  The downside to
such methods is that they can be both computationally expensive and
have large memory requirements. While for some experimental designs
fast iterative methods for the inversion do exist with minimal memory
usage \citep[e.g.,][]{wright1996}, for the more general map-making
problem involving many detectors, things are significantly complicated
both by the need to measure the cross power-spectra for all unique
pairs of bolometers, and then to perform the inversion itself. The
most promising maximum-likelihood method that may one day be applied
to \scuba\ data is ``SANEPIC'', which successfully reduced maps from
the Balloon-borne Large-Aperture Submillimeter Telescope, involving
hundreds of bolometers, and correctly incorporating inter-bolometer
noise correlations \citep[][]{patanchon2008}.

%  Styles of map-making: least
%squares \citep{janssen1992} maximum likelihood
%\citep[e.g.,][]{patanchon2008}; iterative -- see \citet{johnstone2000}
%implementation of \citet{wright1996} to remove chop,
%\citet{kovacs2008} for SHARC-2 etc.; de-correlation using PCA analysis
%\citep[e.g.,][]{laurent2005,scott2008,aguirre2010}.

@@

This paper is organized in the following way. We first describe the
properties of \scuba\ data in Section~\ref{sec:data} to motivate the
data model that we have adopted. Next, the details of the SMURF
algorithm are given in Section~\ref{sec:algorithm}. The paper is
concluded in Section~\ref{sec:examples} with three detailed test cases
that span the majority of observation types likely to be undertaken
with \scuba, with an emphasis on the mitigation of divergence
problems, and characterizing the output maps: (i) Uranus, a bright,
compact source (Section~\ref{sec:point}); (ii) the Lockman Hole, a
blind survey of faint point-like source (Section~\ref{sec:cosmo}); and
(iii) the star-forming region M17, including bright, extended emission
(Section~\ref{sec:extended}). All of the data analyzed in this paper
are publicly available through the CADC \scuba\ raw-data queries page
for the dates and observation numbers given in the
text\footnote{\texttt{http://www3.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/
jcmt/search/scuba2}}. All of the analysis was performed using
\texttt{Starlink} version ... .

@@

In this section we summarize how \scuba\ works,
(Section~\ref{sec:bolos}), give exampels of typical bolometers signals
(Section~\ref{sec:bolosignal}), and the impact of magnetic field
pickup (Section~\ref{sec:magpickup}), and finally illustrate the use
of principal component analysis to explore the correlated noise
properties of \scuba\ data (Section~\ref{sec:pca}).

@@

Incoming light passes through a beam-splitter, and then bandpass
filters, providing simultaneous illumination and wavelength definition
of both the 450 and 850\,\micron\ focal planes. Each focal plan is
populated by four \emph{subarrays} (labeled s4a--s4d at 450\,\micron,
and s8a--s8d at 850\,\micron), each consisting of 32 columns and 40
rows of bolometers. The bolometers are thermal absorbers coupled to
superconducting transition-edge sensors (TESs) for thermometry;
temperature variations in the TESs produce changing currents, and
therefore varying magnetic fields.  The changing magnetic fields are
detected and amplified using chains of superconducting quantum
interference devices (SQUIDs), and the larger output currents are
digitized. Each detector has its own SQUID for the first-stage of the
amplification, but the remaining stages occur within a common chain of
SQUIDs for each of the 32 columns. All 40 rows are read-out in
sequence at a row-visit frequency of about 12\,kHz. Such a high sample
rate is unnecessary to produce maps, so the data are re-sampled to
approximately 200\,Hz before writing to disk.  It should also be noted
that there is an additional 41st row of SQUID readouts that are not
connected to TESs. These \emph{dark SQUIDs} track non-thermal noise
sources that are common to each column's amplifier chain. The
relationship between the output digitized current, $I$, and the input
power, $P$, or $dI/dP$ is established using \emph{flatfield}
observations immediately prior to science observations, in which the
output signal is measured throughout a ramp of the pixel heaters
(which provide a known input power); see Section~2.1 in
\citet{dempsey2012} and Section~5.3 in \citet{holland2012}. Finally,
the conversion to astronomical flux units from power involves a
correction for atmospheric extinction \citep[primarily using the JCMT
Water Vapour Monitor to track line-of-sight opacity variations, see
Section~3 in][]{dempsey2012}, and the application of a flux conversion
factor (FCF) which is established from regular measurements of
calibrators such as Uranus \citep[Section~5 in][]{dempsey2012}.

The predominant noise features in \scuba\ bolometers are: (i) the
effectively white thermal noise produced by the phonon and optical
loading on the absorbers (largely independent from bolometer to
bolometer); and (ii) longer timescale drifts resulting from slow
variations in the mean load (e.g., fridge, telescope, or atmospheric
changes), and also the readout electronics (often correlated amongst
many or all of the bolometers). As such, the power spectra of
bolometer time-series exhibit a ``$1/f^\alpha$'' ($\alpha>0$) shape
below some knee frequency, $f_\mathrm{knee}$, and they are then
relatively flat at higher frequencies. The spectra also include line
features which are predominantly produced by aliased high-frequency
noise sources detected during the multiplexed readout stage as
described above.  Finally, an anti-aliasing filter produces roll-off
in the spectrum at frequencies approaching Nyquist prior to the
200\,Hz re-sampling.

@@

In order to reduce the impact of noise on the final maps, it is
essential that an appropriate scan strategy be used. First, it is
clearly advantageous that the astronomical signals appear in the
lowest-noise region of the power spectrum, namely frequencies above
$f_\mathrm{knee}$. This is achieved by scanning the telescope at large
slew rates, which are practically limited to about
600\,arcsec\,s$^{-2}$ for the JCMT in order to have reasonable
turn-around overheads (the amount of time spent changing direction and
speed, over the largest elevation range possible). At this maximum
speed there is a sample every 3.0\,arcsec, or approximately one third
of the 450\,\micron\ diffraction-limited full-width at half-maximum
(FWHM) -- a typical rule-of-thumb for adequately sampling a Gaussian
point spread function (PSF). The scan \emph{patterns} are designed,
primarily, to mitigate the effects of low-frequency $1/f$ drift. The
simplest scan one might use is a raster ``boustrophedon'' pattern,
i.e., a series of parallel passes across the field. However, such a
scan yields extremely poor performace for structures that are
transverse to the scan direction, since a long time will have elapsed
between each parallel step. In addition, other periodic noise features
may conspire with such a regularly-spaced pattern to produce periodic
structures in the map, or \emph{scan-synchronous} noise.  It has
therefore been recognized that scan strategies that visit each region
of the map on as many different time scales, and scan angles as
possible, offer the best protection against large-scale noise
\citep[see discussion in][and references therein]{kovacs2008b}; such
strategies are said to offer good \emph{cross-linking}. As a tradeoff
between truly random scan patterns and feasibility of implementation
taking into consideration both mechanical and scheduling constraints,
\scuba\ presently has two primary scan modes: PONG, which covers
rectangular areas larger than the array footprint in a closed pattern
that ``bounces'' off the edges at 45\,degree angles (and also usually
combined with a rotation through a number of fixed position angles to
create a ``rotating PONG'' with even better cross-linking) which was
based on a similar scan mode developed at the Caltech Submillimeter
Observatory; and a ``constant velocity daisy'' (CV Daisy) for small
areas (of order the array footprint), in which the telescope moves in
a circle, whose centre also slowly traces out a circle. For a more
complete description of the \scuba\ observing modes, see Section~5 in
\citet{holland2012}.

\begin{figure*}
\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro.pdf}
\includegraphics[width=0.49\linewidth]{bolos_point_mix.pdf}
\caption{A comparison between single bolometer time-series in each of
  the 450 and 850\,\micron\ bands with the mixing chamber temperature
  and azimuth/elevation pointing offsets, before and after upgrading
  the instrument. The grey signals overplotted in the top two panels
  show the residual time-series after removing the common-mode
  signals.  Left: data taken during the S2SRO period, observation 29
  on 20100313. There is a strong correlation between the bolometers
  and the roughly $\sim$25\,s oscillation in the fridge, but only a
  minor correlation with the telescope motion. The nearly flat
  common-mode subtracted signals show: (i) that most of the
  low-frequency signal is common to all of the bolometers; and (ii)
  the non-correlated, and predominantly white noise at 450\,\micron\
  is significantly larger than at 850\,\micron. Right: data taken with
  the fully-commissioned instrument, observation 38 on
  20111112. Unlike the S2SRO data, there is no strong signal produced
  by variations in the fridge temperature. The 450\,\micron\ data have
  minimal correlated low-frequency noise (as evidenced by the lack of
  a difference between the raw and common-mode subtracted signals),
  although the 850\,\micron\ data show a common signal correlated with
  the telescope motion (note that this scan has a significantly larger
  amplitude than the S2SRO data set), mostly likely caused by magnetic
  field pickup (see Section~\ref{sec:magpickup} and
  Fig.~\ref{fig:magpickup}).}
\label{fig:bolos_mix}
\end{figure*}

@@

In Fig.~\ref{fig:bolos_mix} we show sample time-series from single
bolometers in each of the 450 and 850\,\micron\ focal planes, as well
as variations in the mixing chamber temperature (as a proxy for the
focal-plane temperature), and the telescope pointing, for two data
sets, before and after the upgrades that followed S2SRO.

@@

The low-frequency signal component of the S2SRO bolometer output is
also highly correlated amongst bolometers in the same subarray. We
have calculated a common-mode signal, $\mathbf{c}(t)$, as the average
time-series of all the working bolometers. We then fit the amplitude
of $\mathbf{c}(t)$ at each wavelength to the signals shown in
Fig.~\ref{fig:bolos_mix} and remove it, yielding the grey residual
signals. These residuals are quite flat, although still with
noticeable variations. The white noise is also apparent, and larger at
450\,\micron\ as one would expect from the larger backgrounds compared
to 850\,\micron.

@@

Next, we produce power spectral density (PSD) plots for four of the
most sensitive bolometers from both focal planes, using the same two
data sets, in Fig.~\ref{fig:pspec}. To produce this figure, we follow
the convention that the PSD as a function of frequency,
$\mathbf{P}(f)$, is normalized such that the integral over frequency
gives the same variance as the time-series variance across the full
time-series. In other words, given a bolometer signal $\mathbf{b}(t)$,
%
\begin{equation}
\label{eq:psd}
\langle\mathbf{b}^2(t)\rangle = 2 \int_0^{f_\mathrm{N}} \mathbf{P}(f)
df ,
\end{equation}
%
where we only integrate over the positive frequencies up to
$f_\mathrm{N}$, the Nyquist frequency, and the factor of 2 accounts
for the (equal) power that appears at negative frequencies in the
discrete Fourier Transform. The units of the PSD written in this form
are pW$^2$\,Hz$^{-1}$.

@@

An additional noise source that is significant in only a small subset
of the data (and moreso during the S2SRO period) is magnetic field
pickup. Since the bolometer signals are ultimately detected through
the amplification of magnetic fields, any additional changing fields
within the instrument will also be detected.

@@

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{pca.pdf}
\caption{The first twelve components from a principal component
  analysis, ranked by the mean eigenvalues, of combined 450 (s4a
  subarray) and 850\,\micron\ (s8b subarray) time-series bolometer
  data, for the same observation as that used in the right-hand panels
  of Figs.~\ref{fig:bolos_mix} and \ref{fig:pspec}. For each component
  the top plot shows the time-series of the normalized eigenvectors,
  the middle plot its PSD, and the bottom coloured panels the
  eigenvalues for the bolometers at each wavelength (the amplitudes of
  the eigenvectors in the time-series). For reference, both the mean,
  $\bar{\lambda}$, and \rms, $\lambda_\mathrm{rms}$, eigenvalues for
  the bolometers in each subarray are also shown. Most of the
  correlated $1/f$ signal is encapsulated in Components~1 and 2, and
  appears to be dominated by scan-synchronous noise (compare with the
  scan pattern in the right-hand panel of Fig.~\ref{fig:bolos_mix}).
  Components~3, 4, 7 and 11 are transient glitches that only appear in
  a small subset of the bolometers. Component~5 also appears in only a
  few bolometers, and is dominated by a narrow line at 57\,Hz which
  suggests that it is related to the 60\,Hz mains. Components~6 and
  8--10 contain a mixture of broad lines above $\sim$30\,Hz (which are
  probably aliased high-frequency noise), with some low-frequency
  drifts. Finally, the erratic appearance of the eigenvector for
  Component~12 is typical of a poorly-biased detector.}
\label{fig:pca}
\end{figure*}

@@

The majority of the correlated signal at both wavelengths in
Fig.~\ref{fig:pca} is clearly produced by Components~1 and 2. In both
cases, the eigenvector time-series exhibit a roughly periodic signal
that resembles the scan pattern in the right-hand panel of
Fig.~\ref{fig:pca}. While these correlated signals also clearly appear
at both wavelengths, referring to the maps of eigenvalues, they are
considerably stronger at 850\,\micron\ (consistent with the visual
appearance of the bolometer signals in Fig.~\ref{fig:pca}). Also note
that the PSDs for these eigenvectors exhibit nearly a pure $1/f$
signature with almost no line features, or a white-noise plateau
(suggesting that these are high-\snr\ measurements of a purely
low-frequency drift).

@@

Detecting and correcting such a wide variety of steps reliably has
proved to be a challenge. In outline, the following stages are
involved in detecting steps in a single bolometer time-series:

@@

\item median smooth the whole time-series;

\item find the gradient of the median smoothed time-series at each
sample;
